---
title: "Replication: Intergenerational Economic Mobility and the Racial Wealth Gap"
subtitle: "Toney and Robertson (2021), *AEA Papers and Proceedings*"
author: "Oskar Skorge"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    number_sections: true
    code_folding: hide
    fig_width: 8
    fig_height: 5
    df_print: kable
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  message    = FALSE,
  warning    = FALSE,
  fig.align  = "center",
  results    = "asis"
)

# ── Required packages ──────────────────────────────────────────────────────────
# Install any missing packages automatically
required_pkgs <- c(
  "haven",           # read Stata .dta files

"tidyverse",       # data wrangling and plotting
  "fixest",          # panel IV estimation (feols)
  "ivreg",           # classical 2SLS
  "plm",             # panel data models
  "oaxaca",          # Oaxaca-Blinder decomposition
  "modelsummary",    # publication-quality tables
  "kableExtra",      # enhanced kable formatting
  "stargazer",       # regression tables (alternative)
  "gt",              # grammar of tables
  "sandwich",        # robust standard errors
  "lmtest",          # coefficient testing
  "broom",           # tidy model output
  "scales"           # number formatting
)

for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
  library(pkg, character.only = TRUE)
}

# ── Global options ─────────────────────────────────────────────────────────────
options(
  modelsummary_factory_default = "kableExtra",
  scipen = 999
)

# ── Paths ──────────────────────────────────────────────────────────────────────
input_dir  <- "Input/"
output_dir <- "Ouput/"
```

# Introduction

This document replicates the core empirical findings of Toney and Robertson (2021), "Intergenerational Economic Mobility and the Racial Wealth Gap," published in the *American Economic Review: Papers and Proceedings*. The paper examines how parental and grandparental wealth influences the younger generation's income, using data from the Panel Study of Income Dynamics (PSID), 1984--2017.

The replication covers three main sets of results:

1. **Table 1**: Second-stage 2SLS/IV estimates of the effect of stock value (instrumented by extended family wealth) on household income, across parent-child, paternal grandparent-grandchild, and maternal grandparent-grandchild samples.
2. **Table 2**: First-stage results showing the effect of (grand)parental wealth on the younger generation's stock holdings.
3. **Table 3**: Oaxaca-Blinder decomposition of the black-white income gap into endowment, coefficient, and interaction components.

The empirical strategy instruments the log of stock values (stocks and mutual funds held outside of pension wealth) with control variables including (grand)parental wealth, (grand)parental income, educational attainment, and demographic characteristics. The identification rests on the rationale that parental and grandparental wealth indirectly influence offspring income through asset accumulation channels.

---

# Data

## Loading the Datasets

The replication relies on three datasets, each corresponding to a different generational linkage:

```{r load-data}
# ── Parent-Child pairs (Table 1 Col 1, Table 2 Col 1, Table 3 Panel A) ────────
df_parent <- read_dta(paste0(input_dir,
                              "aeapp_mobility_table1_and_summarystatistics.dta"))

# ── Paternal Grandparent-Grandchild pairs (Col 2) ─────────────────────────────
df_paternal <- read_dta(paste0(input_dir,
                                "aeapp_mobility_column2_table1.dta"))

# ── Maternal Grandparent-Grandchild pairs (Col 3) ─────────────────────────────
df_maternal <- read_dta(paste0(input_dir,
                                "aeapp_mobility_column3_table1.dta"))

cat("**Dataset dimensions:**\n\n")
cat(sprintf("- Parent-child: %s observations, %s variables, %s unique children\n",
            format(nrow(df_parent), big.mark = ","),
            ncol(df_parent),
            format(length(unique(df_parent$sib_id)), big.mark = ",")))
cat(sprintf("- Paternal grandparent-grandchild: %s observations, %s variables, %s unique grandchildren\n",
            format(nrow(df_paternal), big.mark = ","),
            ncol(df_paternal),
            format(length(unique(df_paternal$sib_id)), big.mark = ",")))
cat(sprintf("- Maternal grandparent-grandchild: %s observations, %s variables, %s unique grandchildren\n",
            format(nrow(df_maternal), big.mark = ","),
            ncol(df_maternal),
            format(length(unique(df_maternal$sib_id)), big.mark = ",")))
```

## Variable Descriptions

The key variables used across the analysis are defined as follows. All monetary variables are expressed as inverse hyperbolic sine (IHS) transformations.

```{r variable-table}
var_desc <- tibble(
  Variable = c(
    "isib_famincome", "isib_stocks", "ipar_wealth", "ipar_inc",
    "igpar_wealth", "igpar_inc",
    "isib_age", "sib_agesq", "isib_race",
    "isib_educ", "isib_child", "isib_married",
    "isib_employ", "isib_occupation", "sib_pension",
    "isib_2home", "isib_computeruse", "isib_riskpref",
    "sib_sex", "sib_famweight"
  ),
  Description = c(
    "IHS household income of adult child",
    "IHS stock and mutual fund value (outside pension)",
    "IHS parental wealth",
    "IHS parental income",
    "IHS grandparental wealth",
    "IHS grandparental income",
    "Age of adult child",
    "Age squared",
    "Race (1 = non-Latino Black, 0 = non-Latino White)",
    "Educational attainment",
    "Number of children in household",
    "Marital status (1 = married)",
    "Employment status",
    "Professional or managerial occupation (1 = yes)",
    "Pension ownership",
    "Homeownership",
    "Computer/internet use",
    "Risk avoidance",
    "Sex (1 = male)",
    "Family sampling weight"
  ),
  Role = c(
    "Dependent (2nd stage)", "Endogenous (instrumented)", "Instrument/Control", "Control",
    "Instrument/Control", "Control",
    "Control", "Control", "Control",
    "Instrument", "Instrument", "Instrument",
    "Instrument", "Instrument", "Instrument",
    "Instrument", "Instrument", "Instrument",
    "Instrument", "Weight (decomposition)"
  )
)

var_desc %>%
  kbl(caption = "Variable Definitions and Roles in the Empirical Strategy",
      booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  ) %>%
  pack_rows("Outcome and Endogenous Variable", 1, 2) %>%
  pack_rows("Family Background", 3, 6) %>%
  pack_rows("Demographics", 7, 9) %>%
  pack_rows("First-Stage Instruments", 10, 19) %>%
  pack_rows("Weight", 20, 20)
```

---

# Summary Statistics

```{r summary-stats}
# ── Summary statistics for parent-child sample ─────────────────────────────────
sum_vars <- c("isib_famincome", "isib_stocks", "ipar_wealth", "ipar_inc",
              "isib_age", "isib_race", "isib_educ", "isib_child",
              "isib_married", "isib_employ", "isib_occupation", "sib_sex")

sum_stats <- df_parent %>%
  select(all_of(sum_vars)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarise(
    N    = sum(!is.na(Value)),
    Mean = mean(Value, na.rm = TRUE),
    SD   = sd(Value, na.rm = TRUE),
    Min  = min(Value, na.rm = TRUE),
    Med  = median(Value, na.rm = TRUE),
    Max  = max(Value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Variable = factor(Variable, levels = sum_vars)) %>%
  arrange(Variable)

sum_stats %>%
  mutate(across(where(is.numeric) & !N, ~ round(.x, 3))) %>%
  kbl(caption = "Summary Statistics: Parent-Child Sample",
      booktabs = TRUE,
      col.names = c("Variable", "N", "Mean", "Std. Dev.", "Min", "Median", "Max"),
      digits = 3) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  )
```

---

# Empirical Strategy

The identification strategy follows a two-stage least squares (2SLS) instrumental variables approach within a panel data framework. The first stage estimates the effect of extended family wealth on the younger generation's stock holdings:

$$\text{LnStocks}_{it} = C'_{it}\alpha + u_{it} \tag{1}$$

The second stage estimates the causal effect of stock value on household income:

$$\text{LnInc}_{it} = X'_{it}\beta + \text{LnStocks}'_{it}\phi + \varepsilon_{it} \tag{2}$$

where $\text{LnInc}_{it}$ is the IHS-transformed household income for adult child $i$ in survey year $t$, $X'_{it}$ contains age, age squared, and race, and $\text{LnStocks}_{it}$ is instrumented using the $C'_{it}$ vector: (grand)parental wealth, (grand)parental income, educational attainment, number of children, marital status, employment, occupation, pension ownership, homeownership, computer/internet use, risk avoidance, time dummies, and sex.

We employ random effects with EC2SLS estimation and cluster robust standard errors at the individual level, following the original specification.

---

# Replication Results

## Table 1: Second-Stage 2SLS/IV Results {.tabset}

The second stage estimates the effect of (instrumented) log stock value on log household income.

```{r prepare-data}
# ── Helper: create year dummies as numeric columns ─────────────────────────────
# Using factor(year) inside ivreg can fail when some year levels have no
# variation after listwise deletion.  Pre-creating numeric dummies avoids this.

make_year_dummies <- function(df) {
  yrs <- sort(unique(df$year))
  ref  <- yrs[1]                       # first year as reference
  for (y in yrs[-1]) {
    df[[paste0("yr", y)]] <- as.integer(df$year == y)
  }
  df
}

df_parent   <- make_year_dummies(df_parent)
df_paternal <- make_year_dummies(df_paternal)
df_maternal <- make_year_dummies(df_maternal)

# ── Identify year-dummy column names present in each dataset ───────────────────
yr_cols_parent   <- grep("^yr", names(df_parent),   value = TRUE)
yr_cols_paternal <- grep("^yr", names(df_paternal), value = TRUE)
yr_cols_maternal <- grep("^yr", names(df_maternal), value = TRUE)

# ── Filter to complete cases BEFORE estimation ─────────────────────────────────
# This ensures the cluster vector matches the estimation sample exactly.

core_vars <- c("isib_famincome", "isib_stocks", "isib_age", "sib_agesq",
               "isib_race", "isib_educ", "isib_child", "isib_married",
               "isib_employ", "isib_occupation", "sib_pension", "isib_2home",
               "isib_computeruse", "isib_riskpref", "sib_sex")

parent_cc <- df_parent %>%
  filter(complete.cases(across(all_of(c(core_vars, "ipar_wealth", yr_cols_parent)))))

paternal_cc <- df_paternal %>%
  filter(complete.cases(across(all_of(c(core_vars, "igpar_wealth", yr_cols_paternal)))))

maternal_cc <- df_maternal %>%
  filter(complete.cases(across(all_of(c(core_vars, "igpar_wealth", yr_cols_maternal)))))

# ── Drop any year dummies with zero variance in the filtered sample ────────────
drop_constant <- function(df, cols) {
  dominated <- vapply(cols, function(c) var(df[[c]], na.rm = TRUE) == 0, logical(1))
  cols[!dominated]
}

yr_parent   <- drop_constant(parent_cc,   yr_cols_parent)
yr_paternal <- drop_constant(paternal_cc, yr_cols_paternal)
yr_maternal <- drop_constant(maternal_cc, yr_cols_maternal)
```

```{r table1-estimation}
# ── Build formulas dynamically ─────────────────────────────────────────────────
# 2nd-stage: isib_famincome ~ isib_stocks + age + agesq + race
# instruments: all 1st-stage controls + year dummies

build_iv_formula <- function(wealth_var, yr_dummies) {
  inst <- paste(c("isib_age", "sib_agesq", "isib_race",
                   wealth_var, "isib_educ", "isib_child", "isib_married",
                   "isib_employ", "isib_occupation", "sib_pension", "isib_2home",
                   "isib_computeruse", "isib_riskpref", "sib_sex",
                   yr_dummies), collapse = " + ")
  as.formula(paste0(
    "isib_famincome ~ isib_stocks + isib_age + sib_agesq + isib_race | ",
    inst
  ))
}

# ── Column 1: Parent-Child ─────────────────────────────────────────────────────
t1_col1 <- ivreg(build_iv_formula("ipar_wealth", yr_parent),
                  data = parent_cc)

# ── Column 2: Paternal Grandparent-Grandchild ─────────────────────────────────
t1_col2 <- ivreg(build_iv_formula("igpar_wealth", yr_paternal),
                  data = paternal_cc)

# ── Column 3: Maternal Grandparent-Grandchild ─────────────────────────────────
t1_col3 <- ivreg(build_iv_formula("igpar_wealth", yr_maternal),
                  data = maternal_cc)

# ── Cluster-robust standard errors (cluster vector from filtered data) ─────────
vcov1 <- vcovCL(t1_col1, cluster = parent_cc$sib_id)
vcov2 <- vcovCL(t1_col2, cluster = paternal_cc$sib_id)
vcov3 <- vcovCL(t1_col3, cluster = maternal_cc$sib_id)
```

```{r table1-display}
# ── Publication-quality Table 1 ────────────────────────────────────────────────
cm <- c(
  "isib_stocks" = "Log stock value (instrumented)",
  "isib_age"    = "Age",
  "sib_agesq"   = "Age squared",
  "isib_race"   = "Race (Black = 1)",
  "(Intercept)"  = "Constant"
)

gm <- list(
  list("raw" = "nobs", "clean" = "Person-Year Observations", "fmt" = function(x) format(as.integer(x), big.mark = ",")),
  list("raw" = "r.squared", "clean" = "R-squared", "fmt" = 3)
)

# Add number of unique children as a custom row
n_children <- c(
  format(length(unique(parent_cc$sib_id)), big.mark = ","),
  format(length(unique(paternal_cc$sib_id)), big.mark = ","),
  format(length(unique(maternal_cc$sib_id)), big.mark = ",")
)

modelsummary(
  list(
    "(1)\nParent-Child" = t1_col1,
    "(2)\nPaternal GP-GC" = t1_col2,
    "(3)\nMaternal GP-GC" = t1_col3
  ),
  vcov = list(vcov1, vcov2, vcov3),
  coef_map = cm,
  gof_map = gm,
  stars = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
  title = "Table 1: 2SLS/IV Coefficients of Stock Value Influenced by Parental and Grandparental Wealth on the Younger Generation's Income",
  notes = list(
    "Cluster-robust standard errors in parentheses.",
    "Each column reports the second stage results from 2SLS/IV panel regressions.",
    "Log of stock values are instrumented with (grand)parental wealth, income, education, demographics, and time dummies.",
    "Dependent variable: IHS-transformed household income. *** p < 0.01, ** p < 0.05, * p < 0.1."
  ),
  add_rows = data.frame(
    term = "Number of Children",
    V1 = n_children[1],
    V2 = n_children[2],
    V3 = n_children[3]
  ),
  output = "kableExtra"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size = 13
  )
```

### Comparison with Original Table 1

```{r table1-comparison}
# ── Compare replicated coefficients against original paper ─────────────────────
original_t1 <- tibble(
  Sample = c("Parent-Child", "Paternal GP-GC", "Maternal GP-GC"),
  `Original Coef.` = c(0.2523, 0.2200, 0.1826),
  `Original SE`    = c(0.0133, 0.0191, 0.0130),
  `Replicated Coef.` = c(
    coef(t1_col1)["isib_stocks"],
    coef(t1_col2)["isib_stocks"],
    coef(t1_col3)["isib_stocks"]
  ),
  `Replicated SE` = c(
    sqrt(vcov1["isib_stocks", "isib_stocks"]),
    sqrt(vcov2["isib_stocks", "isib_stocks"]),
    sqrt(vcov3["isib_stocks", "isib_stocks"])
  )
) %>%
  mutate(
    `Diff. in Coef.` = `Replicated Coef.` - `Original Coef.`,
    across(where(is.numeric), ~ round(.x, 4))
  )

original_t1 %>%
  kbl(caption = "Table 1 Replication Comparison: Stock Value Coefficient on Household Income",
      booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  ) %>%
  column_spec(6, bold = TRUE, color = "darkblue")
```

*Note:* Minor discrepancies may arise from differences between Stata's `xtivreg` random effects EC2SLS estimator and R's pooled 2SLS approximation via `ivreg`. The Stata implementation accounts for the random effects error structure, whereas this R implementation uses pooled 2SLS with cluster-robust standard errors. Despite this methodological difference, the coefficient magnitudes and significance levels should be qualitatively consistent.

---

## Table 2: First-Stage Results {.tabset}

The first stage regresses log stock value on the full set of instruments and controls, conditional on the second-stage estimation sample.

```{r table2-estimation}
# ── First-stage regressions on the same complete-case samples ──────────────────
# We reuse parent_cc / paternal_cc / maternal_cc and the cleaned year dummies.

build_fs_formula <- function(wealth_var, yr_dummies) {
  rhs <- paste(c("isib_age", "sib_agesq", "isib_race",
                  wealth_var, "isib_educ", "isib_child", "isib_married",
                  "isib_employ", "isib_occupation", "sib_pension", "isib_2home",
                  "isib_computeruse", "isib_riskpref", "sib_sex",
                  yr_dummies), collapse = " + ")
  as.formula(paste0("isib_stocks ~ ", rhs))
}

t2_col1 <- lm(build_fs_formula("ipar_wealth",  yr_parent),   data = parent_cc)
t2_col2 <- lm(build_fs_formula("igpar_wealth", yr_paternal), data = paternal_cc)
t2_col3 <- lm(build_fs_formula("igpar_wealth", yr_maternal), data = maternal_cc)
```

```{r table2-display}
# ── Abbreviated Table 2 (matching paper format: show wealth coef + diagnostics)
cm_t2 <- c(
  "ipar_wealth"  = "Parent's Wealth",
  "igpar_wealth" = "Grandparent's Wealth"
)

gm_t2 <- list(
  list("raw" = "nobs", "clean" = "Person-Year Observations",
       "fmt" = function(x) format(as.integer(x), big.mark = ",")),
  list("raw" = "r.squared", "clean" = "R-squared", "fmt" = 4)
)

n_children_t2 <- c(
  format(length(unique(parent_cc$sib_id)), big.mark = ","),
  format(length(unique(paternal_cc$sib_id)), big.mark = ","),
  format(length(unique(maternal_cc$sib_id)), big.mark = ",")
)

modelsummary(
  list(
    "(1)\nStock Value" = t2_col1,
    "(2)\nStock Value" = t2_col2,
    "(3)\nStock Value" = t2_col3
  ),
  coef_map = cm_t2,
  gof_map = gm_t2,
  stars = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
  title = "Table 2: First Stage Results of Younger Generation's Stock Value Influenced by Parental and Grandparental Wealth",
  notes = list(
    "Standard errors in parentheses.",
    "Each column reports the first stage results from 2SLS/IV panel regressions.",
    "Dependent variable: IHS log of stocks and mutual funds held outside pension wealth.",
    "*** p < 0.01, ** p < 0.05, * p < 0.1."
  ),
  add_rows = data.frame(
    term = "Number of Children",
    V1 = n_children_t2[1],
    V2 = n_children_t2[2],
    V3 = n_children_t2[3]
  ),
  output = "kableExtra"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size = 13
  )
```

### Full First-Stage Results (All Coefficients)

```{r table2-full}
cm_t2_full <- c(
  "isib_age"          = "Age",
  "sib_agesq"         = "Age Squared",
  "isib_race"         = "Race (Black = 1)",
  "ipar_wealth"       = "Parent's Wealth",
  "igpar_wealth"      = "Grandparent's Wealth",
  "isib_educ"         = "Educational Attainment",
  "isib_child"        = "No. Children in HH",
  "isib_married"      = "Married",
  "isib_employ"       = "Employment Status",
  "isib_occupation"   = "Professional/Managerial Occupation",
  "sib_pension"       = "Pension Ownership",
  "isib_2home"        = "Homeownership",
  "isib_computeruse"  = "Computer/Internet Use",
  "isib_riskpref"     = "Risk Avoidance",
  "sib_sex"           = "Sex (Male = 1)",
  "(Intercept)"       = "Constant"
)

modelsummary(
  list(
    "(1)\nParent-Child" = t2_col1,
    "(2)\nPaternal GP-GC" = t2_col2,
    "(3)\nMaternal GP-GC" = t2_col3
  ),
  coef_map = cm_t2_full,
  gof_map = gm_t2,
  stars = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
  title = "Table 2 (Extended): Full First-Stage Regression Results",
  notes = list(
    "Standard errors in parentheses. Year dummies included but not shown.",
    "*** p < 0.01, ** p < 0.05, * p < 0.1."
  ),
  output = "kableExtra"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size = 13
  )
```

### Comparison with Original Table 2

```{r table2-comparison}
original_t2 <- tibble(
  Sample = c("Parent-Child", "Paternal GP-GC", "Maternal GP-GC"),
  `Variable` = c("Parent's Wealth", "Paternal GP Wealth", "Maternal GP Wealth"),
  `Original Coef.` = c(0.0293, 0.0029, 0.0463),
  `Original SE`    = c(0.0066, 0.0143, 0.0136),
  `Original R-sq`  = c(0.1617, 0.1881, 0.1950),
  `Replicated Coef.` = c(
    coef(t2_col1)["ipar_wealth"],
    coef(t2_col2)["igpar_wealth"],
    coef(t2_col3)["igpar_wealth"]
  ),
  `Replicated SE` = c(
    summary(t2_col1)$coefficients["ipar_wealth", "Std. Error"],
    summary(t2_col2)$coefficients["igpar_wealth", "Std. Error"],
    summary(t2_col3)$coefficients["igpar_wealth", "Std. Error"]
  ),
  `Replicated R-sq` = c(
    summary(t2_col1)$r.squared,
    summary(t2_col2)$r.squared,
    summary(t2_col3)$r.squared
  )
) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))

original_t2 %>%
  kbl(caption = "Table 2 Replication Comparison: First-Stage Wealth Coefficients",
      booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  )
```

---

## Table 3: Oaxaca-Blinder Decomposition {.tabset}

The Oaxaca-Blinder decomposition isolates the sources of the black-white income gap into three components: endowments (differences in observed characteristics), coefficients (differences in returns to those characteristics, often interpreted as reflecting structural discrimination), and interaction effects.

### Panel A: Parent-Child Sample

```{r oaxaca-helpers}
# ── Helper: safely extract threefold overall decomposition ─────────────────────
# The oaxaca package may return $threefold$overall as a named vector or matrix.
# This helper converts it to a consistent 3×2 data.frame (coef + se) regardless.

extract_threefold <- function(oax_obj) {
  raw <- oax_obj$threefold$overall

  # Case 1: already a matrix
  if (is.matrix(raw) && nrow(raw) >= 3 && ncol(raw) >= 2) {
    return(data.frame(
      coef = raw[1:3, 1], se = raw[1:3, 2],
      row.names = c("endowments", "coefficients", "interaction")
    ))
  }

  # Case 2: named vector – try pattern matching
  nms <- names(raw)
  if (!is.null(nms)) {
    gv <- function(pat) {
      idx <- grep(pat, nms, ignore.case = TRUE)
      if (length(idx) > 0) raw[idx[1]] else NA_real_
    }
    out <- data.frame(
      coef = c(gv("coef.*endow"), gv("coef.*coeff"), gv("coef.*inter")),
      se   = c(gv("se.*endow"),   gv("se.*coeff"),   gv("se.*inter")),
      row.names = c("endowments", "coefficients", "interaction")
    )
    if (!all(is.na(out$coef))) return(out)
  }

  # Case 3: positional fall-back
  data.frame(
    coef = raw[1:3],
    se   = if (length(raw) >= 6) raw[4:6] else rep(NA_real_, 3),
    row.names = c("endowments", "coefficients", "interaction")
  )
}

# ── Helper: format a single cell as "coef\n(se)" ──────────────────────────────
fmt_coef_se <- function(coef, se) {
  sprintf("%.4f<br>(%.4f)", coef, se)
}
```

```{r table3-estimation}
# ── Prepare 2017 cross-sections ────────────────────────────────────────────────
prep_2017 <- function(df, filter_sample = FALSE) {
  out <- df %>% filter(year == 2017)
  if (filter_sample && "sample" %in% names(out)) out <- out %>% filter(sample == 1)
  out %>%
    mutate(bwrace = case_when(
      isib_race == 0 ~ 0L, isib_race == 1 ~ 1L, TRUE ~ NA_integer_
    )) %>%
    filter(!is.na(bwrace))
}

df_parent_2017   <- prep_2017(df_parent,   filter_sample = TRUE)
df_paternal_2017 <- prep_2017(df_paternal, filter_sample = FALSE)
df_maternal_2017 <- prep_2017(df_maternal, filter_sample = FALSE)

# ── Run threefold Oaxaca-Blinder decompositions ───────────────────────────────

# (1) Parent-Child
oax_parent <- oaxaca(
  isib_famincome ~ isib_educ + ipar_inc + ipar_wealth + isib_child +
    isib_age + sib_agesq + isib_married + isib_employ + sib_sex +
    isib_occupation | bwrace,
  data = df_parent_2017, R = 100
)
oax_A <- extract_threefold(oax_parent)

# (2) Paternal Grandparent-Grandchild
oax_paternal <- oaxaca(
  isib_famincome ~ isib_educ + igpar_inc + igpar_wealth + isib_child +
    isib_age + sib_agesq + isib_married + isib_employ + sib_sex +
    isib_occupation | bwrace,
  data = df_paternal_2017, R = 100
)
oax_B <- extract_threefold(oax_paternal)

# (3) Maternal Grandparent-Grandchild
oax_maternal <- oaxaca(
  isib_famincome ~ isib_educ + igpar_inc + igpar_wealth + isib_child +
    isib_age + sib_agesq + isib_married + isib_employ + sib_sex +
    isib_occupation | bwrace,
  data = df_maternal_2017, R = 100
)
oax_C <- extract_threefold(oax_maternal)

# ── Compute group means ───────────────────────────────────────────────────────
group_means <- function(df) {
  mu_w <- mean(df$isib_famincome[df$bwrace == 0], na.rm = TRUE)
  mu_b <- mean(df$isib_famincome[df$bwrace == 1], na.rm = TRUE)
  c(white = mu_w, black = mu_b, gap = mu_w - mu_b)
}

gm_A <- group_means(df_parent_2017)
gm_B <- group_means(df_paternal_2017)
gm_C <- group_means(df_maternal_2017)
```

```{r table3-combined}
# ── Build combined Table 3 with three columns ─────────────────────────────────

tbl3 <- data.frame(
  Component = c(
    "White Household Income (log)",
    "Black Household Income (log)",
    "White$-$Black Income Gap",
    " ",
    "Endowments",
    "Coefficients",
    "Interaction",
    "  ",
    "Observations"
  ),
  Col1 = c(
    sprintf("%.4f", gm_A["white"]),
    sprintf("%.4f", gm_A["black"]),
    sprintf("%.4f", gm_A["gap"]),
    "",
    fmt_coef_se(oax_A$coef[1], oax_A$se[1]),
    fmt_coef_se(oax_A$coef[2], oax_A$se[2]),
    fmt_coef_se(oax_A$coef[3], oax_A$se[3]),
    "",
    format(nrow(df_parent_2017), big.mark = ",")
  ),
  Col2 = c(
    sprintf("%.4f", gm_B["white"]),
    sprintf("%.4f", gm_B["black"]),
    sprintf("%.4f", gm_B["gap"]),
    "",
    fmt_coef_se(oax_B$coef[1], oax_B$se[1]),
    fmt_coef_se(oax_B$coef[2], oax_B$se[2]),
    fmt_coef_se(oax_B$coef[3], oax_B$se[3]),
    "",
    format(nrow(df_paternal_2017), big.mark = ",")
  ),
  Col3 = c(
    sprintf("%.4f", gm_C["white"]),
    sprintf("%.4f", gm_C["black"]),
    sprintf("%.4f", gm_C["gap"]),
    "",
    fmt_coef_se(oax_C$coef[1], oax_C$se[1]),
    fmt_coef_se(oax_C$coef[2], oax_C$se[2]),
    fmt_coef_se(oax_C$coef[3], oax_C$se[3]),
    "",
    format(nrow(df_maternal_2017), big.mark = ",")
  ),
  stringsAsFactors = FALSE
)

tbl3 %>%
  kbl(
    caption = "Table 3: Oaxaca-Blinder Decomposition of the Black-White Income Gap (2017)",
    col.names = c("", "(1)<br>Parent-Child",
                  "(2)<br>Paternal GP-GC",
                  "(3)<br>Maternal GP-GC"),
    booktabs = TRUE,
    escape  = FALSE,
    align   = c("l", "c", "c", "c")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size  = 13
  ) %>%
  pack_rows("Differential", 1, 3) %>%
  pack_rows("Decomposition", 5, 7) %>%
  footnote(
    general = c(
      "Threefold Oaxaca-Blinder decomposition. Standard errors in parentheses (bootstrapped, R = 100).",
      "White is the reference group. Dependent variable: IHS household income.",
      "Column (1) uses parental income and wealth; columns (2)-(3) use grandparental income and wealth."
    ),
    general_title = "Notes: ",
    footnote_as_chunk = TRUE
  )
```

### Comparison with Original Table 3

```{r table3-comparison}
comparison_t3 <- tibble(
  Panel = c("(1) Parent-Child", "(1) Parent-Child", "(1) Parent-Child",
            "(2) Paternal GP-GC", "(2) Paternal GP-GC", "(2) Paternal GP-GC",
            "(3) Maternal GP-GC", "(3) Maternal GP-GC", "(3) Maternal GP-GC"),
  Component = rep(c("Endowments", "Coefficients", "Interaction"), 3),
  `Original` = c(0.5863, 0.4243, -0.0900,
                  NA, NA, NA,
                  0.4125, 0.1068, 0.0911),
  `Replicated` = c(oax_A$coef[1], oax_A$coef[2], oax_A$coef[3],
                    oax_B$coef[1], oax_B$coef[2], oax_B$coef[3],
                    oax_C$coef[1], oax_C$coef[2], oax_C$coef[3])
) %>%
  mutate(
    Difference = `Replicated` - `Original`,
    across(where(is.numeric), ~ round(.x, 4))
  )

comparison_t3 %>%
  kbl(caption = "Table 3 Replication Comparison",
      booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 13
  ) %>%
  pack_rows("(1) Parent-Child", 1, 3) %>%
  pack_rows("(2) Paternal Grandparent-Grandchild", 4, 6) %>%
  pack_rows("(3) Maternal Grandparent-Grandchild", 7, 9)
```

---

# Density Plot: Black-White Income Distribution

```{r density-plot, fig.cap="Kernel Density of IHS Household Income by Race (2017 Cross-Section, Parent-Child Sample)", fig.width=8, fig.height=2.5}
ggplot(df_parent_2017, aes(x = isib_famincome, fill = factor(isib_race),
                            colour = factor(isib_race))) +
  geom_density(alpha = 0.3, adjust = 1.2) +
  scale_fill_manual(
    values = c("0" = "#2166AC", "1" = "#B2182B"),
    labels = c("White", "Black")
  ) +
  scale_colour_manual(
    values = c("0" = "#2166AC", "1" = "#B2182B"),
    labels = c("White", "Black")
  ) +
  labs(
    x = "IHS Household Income",
    y = "Density",
    fill = "Race",
    colour = "Race"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = c(0.20, 0.70),
    legend.background = element_rect(fill = "white", colour = NA),
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(0, 15, by = 1))
```

---

# Extension: Evaluating and Correcting the Identification Strategy

## Motivation

The causal interpretation of the 2SLS estimates in Toney and Robertson (2021) rests on the exclusion restriction: all variables in the first-stage instrument set $C'_{it}$ that are excluded from the second-stage equation must affect household income *only* through their effect on stock holdings. The second-stage equation controls for age, age squared, and race alone; all remaining variables---including the child's own education, employment status, occupation, marital status, number of children, and sex---are excluded from the structural equation and used solely as instruments.

This exclusion restriction is difficult to defend. The child's educational attainment is a direct determinant of earnings through human capital accumulation, independently of any stock market participation channel. Employment status is mechanically linked to income: employed individuals earn labour income regardless of asset holdings. Marital status affects household income through spousal earnings and household economies of scale. Parental income plausibly influences child income through many non-stock channels, including financing education, providing social capital, and enabling residential sorting.

In this section, we first diagnose the original instrument set by reporting standard IV diagnostics that are absent from the published paper. We then re-specify the 2SLS model with a more defensible partition of variables between the structural equation and the instrument set, and examine whether the core finding survives.

## Step 1: IV Diagnostics for the Original Specification

```{r extension-diagnostics}
# ── IV diagnostics for the original specification (parent-child sample) ────────
# ivreg provides summary diagnostics including weak instrument and Wu-Hausman tests.
# For the overidentification test (Sargan/Hansen J), we use summary(..., diagnostics = TRUE).

diag_orig <- summary(t1_col1, vcov = vcov1, diagnostics = TRUE)

# Extract diagnostic test results
diag_table <- data.frame(
  Test = c(
    "Weak Instruments (first-stage F)",
    "Wu-Hausman (endogeneity)",
    "Sargan (overidentifying restrictions)"
  ),
  Statistic = c(
    sprintf("%.2f", diag_orig$diagnostics["Weak instruments", "statistic"]),
    sprintf("%.2f", diag_orig$diagnostics["Wu-Hausman", "statistic"]),
    sprintf("%.2f", diag_orig$diagnostics["Sargan", "statistic"])
  ),
  `p-value` = c(
    sprintf("%.4f", diag_orig$diagnostics["Weak instruments", "p-value"]),
    sprintf("%.4f", diag_orig$diagnostics["Wu-Hausman", "p-value"]),
    sprintf("%.4f", diag_orig$diagnostics["Sargan", "p-value"])
  ),
  stringsAsFactors = FALSE
)

diag_table %>%
  kbl(caption = "Table 4: IV Diagnostic Tests for Original Specification (Parent-Child Sample)",
      booktabs = TRUE, escape = FALSE,
      col.names = c("Test", "Statistic", "p-value")) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size = 13
  ) %>%
  footnote(
    general = c(
      "Weak instruments: first-stage F-statistic; rule of thumb F > 10 (Stock and Yogo, 2005).",
      "Wu-Hausman: tests endogeneity of the instrumented regressor.",
      "Sargan: tests overidentifying restrictions; rejection implies at least one instrument is invalid."
    ),
    general_title = "Notes: ",
    footnote_as_chunk = TRUE
  )
```

The Sargan test is of particular interest. Under the null hypothesis, all instruments are valid (i.e., uncorrelated with the structural error term). A rejection of this null provides formal statistical evidence that the exclusion restriction is violated---that is, at least one instrument has a direct effect on income beyond the stock-holdings channel. Given the theoretical concerns outlined above, a rejection would be expected.

## Step 2: A More Defensible Specification

We re-partition the variables between the structural equation and the instrument set. The key principle: any variable that has a plausible direct effect on the child's household income should be a control in the second stage, not an excluded instrument.

**Corrected second-stage controls ($X'_{it}$):** age, age squared, race, educational attainment, marital status, employment status, professional/managerial occupation, sex, number of children in household.

**Corrected excluded instruments ($Z_{it}$):** parental wealth, pension ownership, homeownership, computer/internet use, risk avoidance, year dummies.

The rationale for the remaining instruments: parental wealth is the core variable of theoretical interest and plausibly affects child income primarily through enabling asset accumulation (conditional on the child's own human capital and labour market characteristics, which are now controlled for). Pension ownership, homeownership, computer use, and risk avoidance capture financial sophistication and asset-holding propensity that predict stock ownership but---conditional on the child's education, employment, occupation, and marital status---are less likely to directly determine income.

```{r extension-corrected-iv}
# ── Corrected specification: move child characteristics to 2nd stage ───────────

build_iv_corrected <- function(wealth_var, yr_dummies) {
  # Second-stage controls (included instruments)
  x_controls <- c("isib_age", "sib_agesq", "isib_race",
                   "isib_educ", "isib_married", "isib_employ",
                   "isib_occupation", "sib_sex", "isib_child")

  # Excluded instruments (only in first stage)
  z_excluded <- c(wealth_var, "sib_pension", "isib_2home",
                   "isib_computeruse", "isib_riskpref", yr_dummies)

  lhs <- "isib_famincome ~ isib_stocks"
  x_str <- paste(x_controls, collapse = " + ")
  z_str <- paste(c(x_controls, z_excluded), collapse = " + ")

  as.formula(paste0(lhs, " + ", x_str, " | ", z_str))
}

# ── Estimate corrected specification for all three samples ─────────────────────
t4_col1 <- ivreg(build_iv_corrected("ipar_wealth",  yr_parent),   data = parent_cc)
t4_col2 <- ivreg(build_iv_corrected("igpar_wealth", yr_paternal), data = paternal_cc)
t4_col3 <- ivreg(build_iv_corrected("igpar_wealth", yr_maternal), data = maternal_cc)

# ── Cluster-robust standard errors ─────────────────────────────────────────────
vcov4_1 <- vcovCL(t4_col1, cluster = parent_cc$sib_id)
vcov4_2 <- vcovCL(t4_col2, cluster = paternal_cc$sib_id)
vcov4_3 <- vcovCL(t4_col3, cluster = maternal_cc$sib_id)

# ── Diagnostics for corrected spec ─────────────────────────────────────────────
diag_corr <- summary(t4_col1, vcov = vcov4_1, diagnostics = TRUE)
```

```{r extension-comparison-table}
# ── Side-by-side: Original vs Corrected (parent-child sample) ─────────────────

cm_ext <- c(
  "isib_stocks"     = "Log stock value (instrumented)",
  "isib_age"        = "Age",
  "sib_agesq"       = "Age squared",
  "isib_race"       = "Race (Black = 1)",
  "isib_educ"       = "Educational attainment",
  "isib_married"    = "Married",
  "isib_employ"     = "Employed",
  "isib_occupation" = "Professional/managerial",
  "sib_sex"         = "Sex (Male = 1)",
  "isib_child"      = "No. children in HH",
  "(Intercept)"     = "Constant"
)

gm_ext <- list(
  list("raw" = "nobs", "clean" = "Person-Year Observations",
       "fmt" = function(x) format(as.integer(x), big.mark = ","))
)

# Build custom diagnostic rows
diag_rows <- data.frame(
  term = c("First-stage F", "Sargan J (p-value)", "No. excluded instruments"),
  `(1)` = c(
    sprintf("%.1f", diag_orig$diagnostics["Weak instruments", "statistic"]),
    sprintf("%.4f", diag_orig$diagnostics["Sargan", "p-value"]),
    as.character(length(setdiff(
      c("ipar_wealth", "isib_educ", "isib_child", "isib_married",
        "isib_employ", "isib_occupation", "sib_pension", "isib_2home",
        "isib_computeruse", "isib_riskpref", "sib_sex", yr_parent),
      c("isib_age", "sib_agesq", "isib_race"))))
  ),
  `(2)` = c(
    sprintf("%.1f", diag_corr$diagnostics["Weak instruments", "statistic"]),
    sprintf("%.4f", diag_corr$diagnostics["Sargan", "p-value"]),
    as.character(length(c("ipar_wealth", "sib_pension", "isib_2home",
                           "isib_computeruse", "isib_riskpref", yr_parent)))
  ),
  check.names = FALSE
)

modelsummary(
  list(
    "(1)\nOriginal" = t1_col1,
    "(2)\nCorrected" = t4_col1
  ),
  vcov = list(vcov1, vcov4_1),
  coef_map = cm_ext,
  gof_map = gm_ext,
  stars = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
  title = "Table 5: Original vs. Corrected IV Specification (Parent-Child Sample)",
  notes = list(
    "Cluster-robust standard errors in parentheses.",
    "Column (1): original specification where child characteristics serve as excluded instruments.",
    "Column (2): corrected specification where child characteristics are second-stage controls.",
    "Excluded instruments in (2): parental wealth, pension, homeownership, computer use, risk avoidance, year dummies.",
    "*** p < 0.01, ** p < 0.05, * p < 0.1."
  ),
  add_rows = diag_rows,
  output = "kableExtra"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size = 13
  )
```

### Corrected Specification Across All Three Samples

```{r extension-corrected-allcols}
cm_ext_short <- c(
  "isib_stocks"     = "Log stock value (instrumented)",
  "isib_age"        = "Age",
  "sib_agesq"       = "Age squared",
  "isib_race"       = "Race (Black = 1)",
  "isib_educ"       = "Educational attainment",
  "isib_married"    = "Married",
  "isib_employ"     = "Employed",
  "isib_occupation" = "Professional/managerial",
  "sib_sex"         = "Sex (Male = 1)",
  "isib_child"      = "No. children in HH",
  "(Intercept)"     = "Constant"
)

n_children_ext <- c(
  format(length(unique(parent_cc$sib_id)), big.mark = ","),
  format(length(unique(paternal_cc$sib_id)), big.mark = ","),
  format(length(unique(maternal_cc$sib_id)), big.mark = ",")
)

# Diagnostics for columns 2 and 3
diag_corr2 <- summary(t4_col2, vcov = vcov4_2, diagnostics = TRUE)
diag_corr3 <- summary(t4_col3, vcov = vcov4_3, diagnostics = TRUE)

diag_rows_all <- data.frame(
  term = c("Number of Children", "First-stage F", "Sargan J (p-value)"),
  V1 = c(n_children_ext[1],
         sprintf("%.1f", diag_corr$diagnostics["Weak instruments", "statistic"]),
         sprintf("%.4f", diag_corr$diagnostics["Sargan", "p-value"])),
  V2 = c(n_children_ext[2],
         sprintf("%.1f", diag_corr2$diagnostics["Weak instruments", "statistic"]),
         sprintf("%.4f", diag_corr2$diagnostics["Sargan", "p-value"])),
  V3 = c(n_children_ext[3],
         sprintf("%.1f", diag_corr3$diagnostics["Weak instruments", "statistic"]),
         sprintf("%.4f", diag_corr3$diagnostics["Sargan", "p-value"]))
)

modelsummary(
  list(
    "(1)\nParent-Child" = t4_col1,
    "(2)\nPaternal GP-GC" = t4_col2,
    "(3)\nMaternal GP-GC" = t4_col3
  ),
  vcov = list(vcov4_1, vcov4_2, vcov4_3),
  coef_map = cm_ext_short,
  gof_map = gm_ext,
  stars = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
  title = "Table 6: Corrected 2SLS/IV Specification Across All Samples",
  notes = list(
    "Cluster-robust standard errors in parentheses.",
    "Child's education, employment, occupation, marital status, sex, and no. children included as second-stage controls.",
    "Excluded instruments: (grand)parental wealth, pension, homeownership, computer use, risk avoidance, year dummies.",
    "*** p < 0.01, ** p < 0.05, * p < 0.1."
  ),
  add_rows = diag_rows_all,
  output = "kableExtra"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    font_size = 13
  )
```

## Discussion of the Extension

```{r extension-discussion, echo=FALSE}
# Programmatic summary for inline reference
orig_coef  <- round(coef(t1_col1)["isib_stocks"], 4)
corr_coef  <- round(coef(t4_col1)["isib_stocks"], 4)
orig_se    <- round(sqrt(vcov1["isib_stocks", "isib_stocks"]), 4)
corr_se    <- round(sqrt(vcov4_1["isib_stocks", "isib_stocks"]), 4)
sargan_p_orig <- round(diag_orig$diagnostics["Sargan", "p-value"], 4)
sargan_p_corr <- round(diag_corr$diagnostics["Sargan", "p-value"], 4)
fstat_orig <- round(diag_orig$diagnostics["Weak instruments", "statistic"], 1)
fstat_corr <- round(diag_corr$diagnostics["Weak instruments", "statistic"], 1)
```

The diagnostics and re-specification yield three findings.

First, the **Sargan J-test for the original specification** returns a test statistic of `r sprintf("%.1f", diag_orig$diagnostics["Sargan", "statistic"])` with a p-value effectively equal to zero ($p$ `r sprintf("= %.4f", sargan_p_orig)`). This constitutes a decisive rejection of the null that all excluded instruments are valid, providing formal statistical evidence that at least one instrument directly affects income beyond the stock-holdings channel. Given that the excluded set includes the child's own education, employment status, and occupation---all textbook determinants of earnings---this rejection is expected and economically interpretable.

Second, the **first-stage F-statistic** declines from `r fstat_orig` in the original specification to `r fstat_corr` in the corrected specification. This reduction is mechanical: variables such as education and occupation are strong predictors of stock ownership, and moving them to the structural equation reduces the explanatory power of the remaining excluded instruments. Crucially, however, the corrected F-statistic of `r fstat_corr` remains well above the Stock and Yogo (2005) threshold of 10 for the parent-child sample, indicating that the parental wealth and financial-behaviour instruments retain adequate predictive power for stock holdings. The maternal grandparent-grandchild sample (F = `r sprintf("%.1f", diag_corr3$diagnostics["Weak instruments", "statistic"])`) approaches the weak-instrument boundary, suggesting that the three-generation transmission channel is more difficult to identify precisely.

Third, the **point estimate on stock value** falls from `r orig_coef` (SE: `r orig_se`) to `r corr_coef` (SE: `r corr_se`) in the parent-child sample---a reduction of approximately `r sprintf("%.0f", (1 - corr_coef/orig_coef) * 100)` per cent. The direction of this change indicates that the original estimate was upward-biased: the excluded instruments were capturing direct income effects---primarily through education and employment---that were incorrectly attributed to the stock channel. Once we control for these direct determinants in the second stage, the estimated effect of stock holdings on income is smaller but remains statistically significant at the 1\% level. The core insight of Toney and Robertson (2021)---that wealth, channelled through asset accumulation, matters for income---thus survives the correction, albeit at a reduced magnitude.

Fourth, the **race coefficient** is notably attenuated in the corrected specification: it falls from `r round(coef(t1_col1)["isib_race"], 4)` (significant) to `r round(coef(t4_col1)["isib_race"], 4)` (insignificant). In the original specification, race was the only control alongside age, meaning it absorbed all systematic income differences between Black and White respondents that were not mediated by stocks---including education gaps, occupational segregation, employment differentials, and marital status differences. Once these channels are controlled for directly, the residual racial gap conditional on observables shrinks substantially. This finding is consistent with the Oaxaca-Blinder decomposition results in Table 3, which attribute the majority of the black-white income gap to differences in endowments rather than to a pure racial penalty in returns.

A remaining limitation is that the Sargan test continues to reject in the corrected specification ($p$ `r sprintf("= %.4f", sargan_p_corr)`), suggesting that even the retained instruments (pension ownership, homeownership, computer use, risk avoidance) may not be fully exogenous. This points to a deeper identification challenge: isolating the causal effect of a single asset class on income, when the instruments available in household survey data tend to correlate with general financial capability and socioeconomic status, is inherently difficult.

---

# Discussion of Replication

## Methodological Notes

The original paper uses Stata's `xtivreg` command with the `re ec2sls` option, which implements Baltagi's (1981) error components two-stage least squares estimator for panel data with random effects. In this R replication, we approximate this using pooled 2SLS via the `ivreg` package with cluster-robust standard errors at the individual level. While this does not identically reproduce the random effects error component structure, it provides consistent coefficient estimates under the same exclusion restrictions. Differences in standard errors are expected due to this methodological distinction. For the Oaxaca-Blinder decomposition, we use the `oaxaca` R package; minor differences may arise from the handling of sampling weights.

## Summary

The replication confirms the central findings of Toney and Robertson (2021): parental and grandparental wealth have a significant association with offspring income, the transmission extends across three generations, and the black-white income gap is primarily attributable to differences in endowments including family background.

Our extension, however, reveals a material identification concern. The original specification treats the child's education, employment, occupation, marital status, and sex as excluded instruments---requiring that these variables affect income only through stock holdings. The Sargan J-test decisively rejects this restriction. When we re-specify the model with these variables as second-stage controls and restrict the excluded instruments to parental wealth and financial-behaviour proxies, the estimated effect of stock holdings on income falls by approximately `r sprintf("%.0f", (1 - corr_coef/orig_coef) * 100)` per cent, from `r orig_coef` to `r corr_coef`. The finding remains statistically significant, indicating that the wealth-income channel is real but that its magnitude was overstated in the original paper due to invalid exclusion restrictions. The persistent rejection of the Sargan test even in the corrected specification underscores the broader difficulty of achieving clean identification of asset-specific income effects with observational household data.

---

# References

Baltagi, Badi H. 1981. "Simultaneous Equations with Error Components." *Journal of Econometrics*, 17(2): 189--200.

Stock, James H., and Motohiro Yogo. 2005. "Testing for Weak Instruments in Linear IV Regression." In *Identification and Inference for Econometric Models: Essays in Honor of Thomas Rothenberg*, edited by Donald W.K. Andrews and James H. Stock, 80--108. Cambridge University Press.

Toney, Jermaine, and Cassandra L. Robertson. 2021. "Intergenerational Economic Mobility and the Racial Wealth Gap." *American Economic Review: Papers and Proceedings*, 111: 000--000.

---

*Replication conducted using R version `r R.version.string`. All code is embedded and reproducible.*

```{r session-info, echo=TRUE}
sessionInfo()
```
